---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# harsatextras

<!-- badges: start -->
<!-- badges: end -->

The goal of harsatextras is to add some extra functions that extends the [HARSAT](https://osparcomm.github.io/HARSAT/) package. For the general HARSAT user, 
the most useful functions are probably the functions for extracting plot data for trend plots (`get_assessment_data()`), for making a 
ggplot2 trend plot (`ggplot_assessment()`), and for opening a simple Shiny app for viewing trend plots (`open_assessment_app()`. There are also some functions that
are slightly modified versions of HARSAT functions, designed for using HARSAT in a [targets](https://docs.ropensci.org/targets/) pipeline (especially using static branching). 


## Installation

You can install the development version of harsatextras from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DagHjermann/harsatextras")
```

## Example: extracting the data needed for making a time series plot  

This is a basic example which shows you how to extract the plot data from an assessment object 
(i.e. the output of `harsat::run_assessment()`). We will here use the built-in object
`assessment_part1`.  

```{r get_assessment_data1}

library(harsatextras)
plotdata <- get_assessment_data(assessment_part1)

```
The `plotdata` object contains a list with one item per time series:

```{r get_assessment_data2}
str(plotdata, 1)
```

Each list item is a list containing `data` (the raw data), `assessment` (also a list, containing the results of the assessment),
the metadata for the series (`series`) and for the whole data set (`info`), and the series name (`output_id`). For instance, here is 
the contents for time series number 1:

```{r get_assessment_data3}
i <- 1
str(plotdata[[i]], 1)
```

We can plot the raw data using the following command (for simplicity, we ignore the 'cencoring' column,
thereby ignoring to separate between data over LOQ and data under LOQ):

```{r plot_assessment_data1}

plot(concentration ~ year, plotdata[[i]]$data, 
     main = plotdata[[i]]$output_id)


```

The data needed for making a trend plot are found in `assessment`, in the `pred` item. In this case (as
for most chemical parameters), the y axis reflects that the trend assessment has been performed 
on log-transformed data: 

```{r plot_assessment_data2}

plot(fit ~ year, plotdata[[i]]$assessment$pred, 
     main = plotdata[[i]]$output_id,
     type = "l")
lines(ci.lower ~ year, plotdata[[i]]$assessment$pred, 
     type = "l", lty = "dashed")
lines(ci.upper ~ year, plotdata[[i]]$assessment$pred, 
     type = "l", lty = "dashed")

```
The statistical distribution used for assessment data can be found here:

```{r plot_assessment_data3}

plotdata[[1]]$series$distribution

```
The following is an example plot of both raw data and trend lines using ggplot2:

```{r, plot_assessment_data4}

library(ggplot2)

ggplot(plotdata[[i]]$assessment$fullData, aes(year)) +
  geom_ribbon(
    data = plotdata[[i]]$assessment$pred,
    aes(ymin = exp(ci.lower), ymax = exp(ci.upper)),  # note; hard-coded exp
    fill = "lightblue") +
  geom_path(
    data = plotdata[[i]]$assessment$pred,
    aes(y = exp(fit))) +
  geom_point(
    aes(y = concentration, color = censoring),
    color = "darkred") +
  scale_y_log10() +
  labs(title = plotdata[[i]]$output_id)

```

`ggplot_assessment()` is a simple function for making trend plots using the ggplot2 package. See
?ggplot_assessment for options.

```{r, ggplot_assessment}

ggplot_assessment(plotdata[[i]])

```


